\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-07-09 paper.qmd
revised start of application section

In the field of single-cell studies, a common analytical task involves clustering to identify groups of cells with similar expression profiles. Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories. To illustrate the importance of NLDR techniques and parameter selection in identifying clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset [@chen2023] is used. In a study by @chen2023, this dataset was used to demonstrate how UMAP represents clusters (see @fig-umap-author). As shown in @fig-umap-author, there are three distant and well-separated clusters.

@@

<!-- UMAP parameters suggested by the author (n_neighbors = 30, min_dist = 0.3)-->
```{r}
#| fig-cap: "$2\\text{-}D$ layout from UMAP applied for the PBMC3k dataset. Is this a best representation of the original data? The parameter setting is n_neighbors = 30, min_dist = 0.3."
#| label: fig-umap-author
#| out-height: 30%
#| fig-pos: H

>>>>

In the field of single-cell studies, a common analytical task involves clustering to identify groups of cells with similar expression profiles. NLDR is commonly used to display clusters, and help to verify the results. For example, @chen2023 illustrates the use of UMAP to identify clusters in Human Peripheral Blood Mononuclear Cells (PBMC3k). @fig-umap-author is a reproduction of the published plot. It shows three well-separated clusters.

@@

<!-- UMAP parameters suggested by the author (n_neighbors = 30, min_dist = 0.3)-->
```{r}
#| fig-cap: "Reproduction of plot published in Chen et al. (2023) showing a $2\\text{-}D$ layout from UMAP applied for the PBMC3k dataset. The parameter settings, beyond the defaults, are n_neighbors = 30 and min_dist = 0.3. The question is whether this is an accurate representation of structure in the high-dimensional data."
#| label: fig-umap-author
#| out-height: 30%
#| fig-pos: H

### 2024-07-09 paper.tex
revised start of application section

In the field of single-cell studies, a common analytical task involves
clustering to identify groups of cells with similar expression profiles.
Analysts often turn to NLDR techniques to verify and identify these
clusters and explore developmental trajectories. To illustrate the
importance of NLDR techniques and parameter selection in identifying
clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset
\citep{chen2023} is used. In a study by \citet{chen2023}, this dataset
was used to demonstrate how UMAP represents clusters (see
Figure~\ref{fig-umap-author}). As shown in Figure~\ref{fig-umap-author},
there are three distant and well-separated clusters.

@@

\caption{\label{fig-umap-author}\(2\text{-}D\) layout from UMAP applied
for the PBMC3k dataset. Is this a best representation of the original
data? The parameter setting is n\_neighbors = 30, min\_dist = 0.3.}

>>>>

In the field of single-cell studies, a common analytical task involves
clustering to identify groups of cells with similar expression profiles.
NLDR is commonly used to display clusters, and help to verify the
results. For example, \citet{chen2023} illustrates the use of UMAP to
identify clusters in Human Peripheral Blood Mononuclear Cells (PBMC3k).
Figure~\ref{fig-umap-author} is a reproduction of the published plot. It
shows three well-separated clusters.

@@

\caption{\label{fig-umap-author}Reproduction of plot published in Chen
et al.~(2023) showing a \(2\text{-}D\) layout from UMAP applied for the
PBMC3k dataset. The parameter settings, beyond the defaults, are
n\_neighbors = 30 and min\_dist = 0.3. The question is whether this is
an accurate representation of structure in the high-dimensional data.}
